---
title: "Electricity Report"
author: "unaltugba"
date: "11&11&2020"
output: html_document
---

## About Data

This dataset consists of six different values in September 2020:

- Date
- MCP: Market clearing price 
- SMP: System Marginal Price 
- Positive Imbalance Price: min(PTF,SMF)*0,97
- Negative Imbalance Price: max(PTF,SMF)*1,03
- SMP Direction : Energy Surplus or Energy Deficit

You can find the data [here](https://rapor.epias.com.tr/rapor/xhtml/ptfSmfListeleme.xhtml)

## Data Analysis

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(readxl)
library(lubridate)
```

```{r}
energy_data<-read_excel("C:/Users/unalt/Desktop/MEF/Data Analytics Essentials/week3/ptf-smf.xlsx")
summary(energy_data)
```
### Data Manipulations

```{r}
new_energy_data<- energy_data %>%
  transform(Date = as.POSIXct(Date, format='%d.%m.%y %H:%M')) %>%
  mutate(New_Date= as.Date(Date)) %>% 
  mutate(Day = weekdays(Date)) %>% mutate(SMP_MCP_Diff=SMP-MCP) %>%
  mutate(Time= format(energy_data$Date,"%H:%M:%S")) %>%
  mutate(Hour=lubridate::hour(Date))

summary(new_energy_data)
```
### Data Visualization

```{r}
geom_graph <- new_energy_data %>% group_by(Hour) %>% 
  summarise(Avg_MCP=mean(MCP), Min_MCP=min(MCP), Max_MCP=max(MCP), Avg_SMP=mean(SMP),  
            Min_SMP=min(SMP), Max_SMP=max(SMP))
```

```{r}
hourly_mcp<-ggplot(geom_graph, aes(x=Hour)) + geom_line(aes(y=Avg_MCP, color="Average Hourly MCP"), size=0.8) +
              geom_line(aes(y=Min_MCP, color="Minimum Hourly MCP"), size=0.8) + 
              geom_line(aes(y=Max_MCP, color="Maximum Hourly MCP"), size=0.8) +
              labs(x="Hour", y="MCP (TL)") + theme_minimal() +
              ggtitle("Hourly MCP Trend")

hourly_mcp + theme(
  plot.title = element_text(color="darkred", size=14, face="bold"))
```

Although there are deviations in the MCP value during the day, considering the average MCP value, it is seen that the most energy need occurs between 1 pm and 4 pm.


```{r}
hourly_smp<-ggplot(geom_graph, aes(x=Hour)) + geom_line(aes(y=Avg_SMP, color="Average Hourly SMP"), size=0.8) +
              geom_line(aes(y=Min_SMP, color="Minimum Hourly SMP"), size=0.8) + 
              geom_line(aes(y=Max_SMP, color="Maximum Hourly SMP"), size=0.8) +
              labs(x="Hour", y="SMP (TL)") + theme_minimal() +
              ggtitle("Hourly SMP Trend")

hourly_smp + theme(
  plot.title = element_text(color="darkred", size=14, face="bold"))
```

It is seen that the SMP values and the MCP values given in the previous graph move in parallel. Especially between 4 pm and 5 pm, it is seen that the energy requirement is approximately 4 times the normal energy requirement.


```{r, fig.width=10, fig.height=7}
cor<- ggplot(new_energy_data, aes(MCP, SMP, color=Day)) + geom_point() + geom_smooth(method= 'lm') + 
        theme_classic() + ggtitle("Day-Based Correlation Between SMP and MCP")

cor + theme(
  plot.title = element_text(color="darkred", size=14, face="bold"))
```

Looking at the day-based of SMP and MCP data, as expected, it was observed that the energy need at the weekend is lower than during the weekday. The highest energy requirement occurred on Thursday


```{r}
col_graphs<- new_energy_data %>% 
  group_by(Date) %>% select(Date, SMP_MCP_Diff, SMPDirection)


diff_smp_mcp<-ggplot(col_graphs, aes(x=Date, y=SMP_MCP_Diff, fill=SMPDirection)) + geom_col() + 
                scale_fill_manual("SMPDirection", values=c("#FF9900","#CC0000","#009E73")) +
                labs(x="Date", y="SMP - MCP") + theme_light() +
                ggtitle("Energy Deficit or Surplus Trend", "September 2020 Electricity Market Prices")
  
diff_smp_mcp + theme(
  plot.title = element_text(color="darkred", size=14, face="bold"),
  plot.subtitle = element_text(color="darkred", size=11))

```

Looking at the general data for September, it is seen that the balance of production and consumption has never been achieved and the days of energy deficit are the majority

