---
title: "Week3-BonusExercise"
author: "unaltugba"
date: "04 11 2020"
output: html_document
---


```{r df}
library(readxl)

df<- read_xlsx("/Users/unalt/Desktop/MEF/Data Analytics Essentials/week3/EVDS_istanbul_property_data.xlsx",skip=36,n_max=94,
               col_names=c('Date','TotalSales','EncumberedSales','FirsthandSales',
                           'SecondhandSales','ForeignSales','NewHousePriceIndex',
                           'HouseLevel','HouseUnitPrice'))

library(ggplot2)

p<- ggplot(df, aes(Date, TotalSales, color=NewHousePriceIndex)) + 
      geom_point(aes(size=HouseUnitPrice)) +
      ggtitle("Istanbul Property Trend") +
      xlab("Date") + ylab("Total House Sales")
```


```{r, echo=FALSE}
p + theme(
  plot.title = element_text(color="tomato2", size=14, face="bold.italic", hjust=0.5),
  axis.title.x = element_text(color="dimgrey", size=12, face="bold"),
  axis.title.y = element_text(color="dimgrey", size=12, face="bold"),
  axis.text.x = element_text(angle=90, size=6)
)

```


### When the real estate data of Istanbul between 2013-2020 were examined, the following findings were obtained:

- It is seen that the Istanbul new housing price index has started to rise since July 2016.
- Istanbul housing unit prices have been on the rise since March 2015.
- Despite the increase in new housing price index and housing unit prices, house sales in Istanbul have been on an rising trend since September 2019.

